### 2023-06-26 paper/main.tex
LWFS based on modal band-limitation
...

By investigating the sampling process in the frequency-wavenumber domain, the spectrum of the sampled driving function is given by
\begin{equation}
    \tilde{D}_s(k_s,\omega) = \sum_{\nu = -\infty}^{\infty} \tilde{D}\left(k_s- \nu \frac{2\pi}{\Delta s},\omega\right) \Delta s,
    \label{eq:sampling_fd}
\end{equation}
where $t \rightarrow \omega$ and $s \rightarrow k_s$ are Fourier-transform pairs.
Obviously, for a closed SSD contour the continuous spatial Fourier-transform yields a discrete spectrum, consisting of non-zero components only on the multiples of the spatial period length, i.e. at
\begin{equation}
    k_s = m \frac{ 2 \pi }{R_0},
\end{equation}
constituting a Fourier series.

In the followings the spatial spectrum refers to the distribution of the non-zero spectral/wavenumber values, and the corresponding wavenumber refers to the spatial frequency of the harmonic basis functions, measured along the closed curve.
Figure \ref{Fig:aliasing} (c) and (d) illustrate the quasi-continuous and the sampled driving function's spectra.

@@

\section{Time domain antialiasing filter design}

@@

Assume an arbitrary filter impulse response defined in the spatial domain, denoted by $h_x(s)$.
Again, $s$ is a suitable parametrization of the SSD contour.
The spatially filtered driving function is obtained from the convolution of the conventional driving functions and the filter impulse response along the SSD contour
\begin{equation}
    d'_x(s,t) = h_x(s) \ast_x d(s,t) = \int h_x(s-s_0) \, d(s_0,t) \, \td s_0,
    \label{eq:spatial_conv}
\end{equation}
where $\ast_x$ denotes a circular convolution for a convex SSD contour or a linear convolution in case of an infinite long linear SSD.

@@

The temporal convolution in \eqref{eq:temp_conv} can be evaluated by substituting the general WFS driving function \eqref{eq:d_wfs_td} and exploiting the sifting property of the Dirac-delta, yielding
\begin{align}
    d'_t(s,t) = A^D(s) \, h_t \!\left(s,t- \phi^P(s) \right) \\
    d'_t(s,t + \phi^P(s) ) = A^D(s) \, h_t \!\left(s,t\right).
    \label{eq:temp_conv_eval}
\end{align}
In the second equation the time shift of $\phi^P(s)$ was applied for the sake of simplicity.

The shifted spatially filtered driving function reads as
\begin{multline}
    d'_x(s,t+ \phi^P(s) ) = \int h_x(s-s_0) \, d(s_0,t+ \phi^P(s) ) \, \td s_0 = \\
    \int h_x(s-s_0) \, A^D(s_0)\, \delta(s_0,t+ \left(\phi^P(s)-\phi^P(s_0)\right)  ) \, \td s_0.
    \label{eq:spat_conv2}
\end{multline}
In order to evaluate the spatial convolution the generalized sifting property of the Dirac-delta may be applied \cite{Firtha2019phd}, which states that
\begin{equation}
    \int f(s_0) \delta(g(s_0)) \td s_0 = \sum_i \frac{f(s_i)}{| \frac{\partial}{\partial s_0} g(s_0) |_{s = s_i}}, \hspace{3mm} g(s_i) = 0.
\end{equation}
In the present problem the zeros of the Dirac's argument are found where
\begin{equation}
    \phi^P(s_i) = t + \phi^P(s)
    \label{eq:zero}
\end{equation}
is satisfied.
In the following it is assumed that a single zero $s_i$ exists, satisfying \eqref{eq:zero}.
This assumption means that the wavefront arrives at each SSD element at a unique time instant, being strictly true for a virtual plane wave and a linear SSD.
The derivative of the phase function is given by
\begin{equation}
    \frac{\partial}{\partial s_0} g(s_0) = - \frac{\partial}{\partial s_0} \phi^P(s_0) = -\frac{\hat{k}_t^P(s_0)}{c},
\end{equation}
i.e. it is the tangential component of the local propagation vector.
With the above assumptions the integral \eqref{eq:spat_conv2} can be evaluated, yielding
\begin{equation}
    d'_x(s,t+ \phi^P(s) ) = h_x(s-s_i) \, A^D(s_i)
    .
    \label{eq:spat_conv_eval}
\end{equation}

@@

\subsubsection{Local plane wave approximation}
The above formulation already allows one to transform an arbitrary spatial impulse response into an equivalent temporal filter for each SSD element, as long as the virtual field model is known and \eqref{eq:zero} can be solved.
In order to give a more general solution it is assumed that the virtual field is locally plane, being a usual high frequency assumption in WFS theory.
As a further general WFS assumption it is supposed that the SSD is locally plane.
These requirements inherently ensure that a single solution exists for \eqref{eq:zero}\footnote{Except for a plane wave arriving normally to the SSD, at which case the temporal filter is transformed into a Dirac-delta.}.
With these assumptions the phase function is given as
\begin{equation}
    \phi^P(\vx) = \hat{k}^P_x(\vx)x +  \hat{k}^P_y(\vx) y
\end{equation}
and \eqref{eq:zero} is satisfied where
\begin{equation}
    s_i = s + \frac{c \cdot t}{\hat{k}_t^P(s)}
    .
\end{equation}
Finally, as a crucial approximation it is assumed that both the amplitude of the driving function and the propagation vector varies slowly along the SSD, i.e $A^D(s_i) \approx A^D(s)$ and $\hat{k}_t^P(s_i) \approx \hat{k}_t^P(s)$ holds.
With these assumptions the corresponding filter transform reads as
\begin{equation}
    h_t\left(s,t\right) =\frac{ c }{|\hat{k}_t^P(s)|} h_x\left(-t \,\frac{c}{\hat{k}_t^P(s)}\right).
    \label{eq:temporal_transform}
\end{equation}

@@

\subsection{Application for Local Wave Field Synthesis}

@@

The result of antialiasing filtering is depicted in Figure \ref{Fig:symm_antialiasing} for the synthesis scenario discussed in the foregoing.
Figure (a) shows the result of the synthesis, while Figure (b) illustrates the spectrum of the bandlimited, sampled driving functions.
\begin{figure}[h!]
    \begin{center}
        \begin{overpic}[width = 0.75\columnwidth]{figs/antialiased_synthesis.png}
            \footnotesize \put(0,2){(a)}
        \end{overpic}
        \\
        \begin{overpic}[width = 0.75\columnwidth]{figs/antialiased_synthesis_spectrum.png}
            \footnotesize \put(0,0){(b)}
        \end{overpic}
    \end{center}
    \caption{Symmetric antialiasing filtering}
    \label{Fig:symm_antialiasing}
\end{figure}
It is highlighted, that the presented antialiasing strategy ensures highly suppressed aliasing waves in the center of the circular source array.
In other positions of the listening area lateral aliasing wavefronts remain in the reproduced wave field, while the target, virtual wave front is bandlimited.
This phenomena can be easily understood by investigating the synthesis scenario in a geometrical manner:

>>>>

By investigating the sampling process in the frequency-wavenumber domain, the spectrum of the sampled driving function is given by
\begin{equation}
    \tilde{D}_s(k_s,\omega) = \sum_{\nu = -\infty}^{\infty} \tilde{D}\left(k_s- \nu \frac{2\pi}{\Delta s},\omega\right) \Delta s,
    \label{eq:sampling_fd}
\end{equation}
where $t \rightarrow \omega$ and $s \rightarrow k_s$ are Fourier-transform pairs.
Obviously, for a closed SSD contour the continuous spatial Fourier-transform yields a discrete spectrum, consisting of non-zero components only on the multiples of the spatial wavenumber, i.e. at
\begin{equation}
    k_s = \frac{m}{R_0},
\end{equation}
constituting a Fourier series.

In the followings the spatial spectrum refers to the distribution of the non-zero wavenumber values, and the corresponding wavenumber refers to the spatial frequency of the harmonic basis functions, measured along the closed curve.
Figure \ref{Fig:aliasing} (c) and (d) illustrate the quasi-continuous and the sampled driving function's spectra.

@@

\section{Time domain antialiasing filter}

@@

Assume an arbitrary filter impulse response defined in the spatial domain, denoted by $h_x(s)$.
Again, $s$ is a suitable parametrization of the SSD contour.
The spatially filtered driving function is obtained from the convolution of the conventional driving functions and the filter impulse response along the SSD contour
\begin{equation}
    d'_x(s,t) = h_x(s) \ast_s d(s,t) = \int h_x(s-s_0) \, d(s_0,t) \, \td s_0,
    \label{eq:spatial_conv}
\end{equation}
where $\ast_x$ denotes a circular convolution for a convex SSD contour or a linear convolution in case of an infinite long linear SSD.

@@

The temporal convolution in \eqref{eq:temp_conv} can be evaluated by substituting the general WFS driving function \eqref{eq:d_wfs_td} and exploiting the sifting property of the Dirac-delta, yielding
\begin{align}
    d'_t(s,t) = A^D(s) \, h_t \!\left(s,t- \tau(s) \right) \\
    d'_t(s,t + \tau(s) ) = A^D(s) \, h_t \!\left(s,t\right).
    \label{eq:temp_conv_eval}
\end{align}
In the second equation the time shift of $\tau(s)$ was applied for the sake of simplicity.

The shifted spatially filtered driving function reads as
\begin{multline}
    d'_x(s,t+ \tau(s) ) = \int h_x(s-s_0) \, d(s_0,t+ \tau(s) ) \, \td s_0 = \\
    \int h_x(s-s_0) \, A^D(s_0)\, \delta(s_0,t+ \left(\tau(s)-\tau(s_0)\right)  ) \, \td s_0.
    \label{eq:spat_conv2}
\end{multline}
In order to evaluate the spatial convolution the generalized sifting property of the Dirac-delta may be applied \cite{Firtha2019phd}, which states that
\begin{equation}
    \int f(s_0) \delta(g(s_0)) \td s_0 = \sum_i \frac{f(s_i)}{| \frac{\partial}{\partial s_0} g(s_0) |_{s = s_i}}, \hspace{3mm} g(s_i) = 0.
\end{equation}
In the present problem the zeros of the Dirac's argument are found where
\begin{equation}
    \tau(s_i) = t + \tau(s)
    \label{eq:zero}
\end{equation}
is satisfied.
In the following it is assumed that a single zero $s_i$ exists, satisfying \eqref{eq:zero}.
This assumption means that the wavefront arrives at each SSD element at a unique time instant, being strictly true for a virtual plane wave and a linear SSD.
The derivative of the phase function is given by
\begin{equation}
    \frac{\partial}{\partial s_0} g(s_0) = - \frac{\partial}{\partial s_0} \tau(s_0) = -\frac{\hat{k}_t^P(s_0)}{c},
\end{equation}
i.e. it is the tangential component of the local propagation vector.
With the above assumptions the integral \eqref{eq:spat_conv2} can be evaluated, yielding
\begin{equation}
    d'_x(s,t+ \tau(s) ) = h_x(s-s_i) \, A^D(s_i)
    .
    \label{eq:spat_conv_eval}
\end{equation}

@@

\subsubsection{Local plane wave approximation}
The above formulation already allows one to transform an arbitrary spatial impulse response into an equivalent temporal filter for each SSD element, as long as the virtual field model is known and \eqref{eq:zero} can be solved.
In order to give a more general solution it is assumed that the virtual field is locally plane, being a usual high frequency assumption in WFS theory.
As a further general WFS assumption it is supposed that the SSD is locally plane.
These requirements inherently ensure that a single solution exists for \eqref{eq:zero}\footnote{Except for a plane wave arriving normally to the SSD, at which case the temporal filter is transformed into a Dirac-delta.}.
With these assumptions the phase function is given as
\begin{equation}
    \tau(\vx) = \hat{k}^P_x(\vx)x +  \hat{k}^P_y(\vx) y
\end{equation}
and \eqref{eq:zero} is satisfied where
\begin{equation}
    s_i = s + \frac{c \cdot t}{\hat{k}_t^P(s)}
    .
\end{equation}
Finally, as a crucial approximation it is assumed that both the amplitude of the driving function and the propagation vector varies slowly along the SSD, i.e $A^D(s_i) \approx A^D(s)$ and $\hat{k}_t^P(s_i) \approx \hat{k}_t^P(s)$ holds.
With these assumptions the corresponding filter transform reads as
\begin{equation}
    h_t\left(s,t\right) =\frac{ c }{|\hat{k}_t^P(s)|} h_x\left(-t \,\frac{c}{\hat{k}_t^P(s)}\right).
    \label{eq:temporal_transform}
\end{equation}

@@

\subsection{Application for antialiasing filtering}

@@

The result of antialiasing filtering is depicted in Figure \ref{Fig:symm_antialiasing} for the synthesis scenario discussed in the foregoing.
In the present case the sampling arc length is given by $\Delta s = \frac{2 \pi R_0}{N}$, i.e. the space-dependent cut-off frequency reads as\begin{equation}
    \omega_c(s) = \frac{ N}{ 2 R_0 } \frac{c}{\hat{k}_t^P(s)}.
\end{equation}
The antialiasing filters were a 4-th order Butterworth deisgn.

Figure (a) shows the result of the synthesis, while Figure (b) illustrates the spectrum of the bandlimited, sampled driving functions.
\begin{figure}[h!]
    \begin{center}
        \begin{overpic}[width = 0.75\columnwidth]{figs/antialiased_synthesis.png}
            \footnotesize \put(0,2){(a)}
        \end{overpic}
        \\
        \begin{overpic}[width = 0.75\columnwidth]{figs/antialiased_synthesis_spectrum.png}
            \footnotesize \put(0,0){(b)}
        \end{overpic}
    \end{center}
    \caption{Symmetric antialiasing filtering}
    \label{Fig:symm_antialiasing}
\end{figure}
It is highlighted, that the presented antialiasing strategy ensures highly suppressed aliasing waves in the center of the circular source array.
In other positions of the listening area lateral aliasing wavefronts remain in the reproduced wave field, while the target, virtual wave front is bandlimited.
This phenomena can be easily understood by investigating the synthesis scenario in a geometrical manner:
